Priors

### 2023-06-20 reports/county_correlations.Rmd
update rmarkdown files

---
title: "County Correlations"
author: "Quinn White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

>>>>

---
title: "County Results"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)

```

```{r theme_c}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text = element_text(              #axis text
                   size = 9),
      legend.text = element_text(size = 12),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 11, face = "bold", color="white"),
      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

```{r load-data}

targets_dir <- here("_targets_6_19")

county <- tar_read(ma_v1, 
                   store =targets_dir) %>% 
  mutate(version = "v1") %>%
  bind_rows(
    tar_read(ma_v2,
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(ma_v3, 
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(ma_v4, 
             store =targets_dir)
  # )%>%
  #  bind_rows(
  #   tar_read(ma_v5, 
  #            store =targets_dir)
  # )%>%
  #  bind_rows(
  #   tar_read(ma_v6, 
  #            store =targets_dir)
   )

waste <- tar_read(waste, store =targets_dir)
covidestim <- tar_read(covidestim_biweekly_county,
                        store =targets_dir)

dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))

```

```{r}

wjoined <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
  inner_join(waste) %>%
  left_join(covidestim, relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup()

counties_with_waste <- wjoined %>%
  group_by(fips) %>%
  mutate(obs_w_notna = sum(!is.na(mean_conc))) %>%
  filter(obs_w_notna != 0) %>%
  pull(fips) %>%
  unique()

versions <- tibble(
  version = c("v1", "v2", "v3", "v4", "v5", "v6"),
  vlabel = factor(
    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "$\\beta$ Centered at Emp. Value",
    "$P(S_1|untested)$ Centered at Emp. Value",
    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
    ),
    levels = c(
        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
        "$\\beta$ Centered at Emp. Value",
         "$P(S_1|untested)$ Centered at Emp. Value",
        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"

    ) )
)

wjoined <- wjoined %>%
  left_join(versions)

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(mean_conc_std = mean_conc/ max(mean_conc),
         exp_cases_median_std = exp_cases_median / max(exp_cases_median),
         exp_cases_lb_std = exp_cases_lb /max(exp_cases_median),
         exp_cases_ub_std = exp_cases_ub / max(exp_cases_median),
         infections_std  = infections/max(infections),
         positive_std = positive/max(positive)) %>%
  ungroup()

```

# Correlations

## Normalization Strategy 1: By Values at Biweek Where Wastewater Concentration is Maximized

```{r}

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(mean_conc_std = mean_conc/ max(mean_conc),
         max_conc = max(mean_conc),
         exp_cases_median_std = exp_cases_median / exp_cases_median[which.max(mean_conc)],
         exp_cases_lb_std = exp_cases_lb /exp_cases_median[which.max(mean_conc)],
         exp_cases_ub_std = exp_cases_ub / exp_cases_median[which.max(mean_conc)],
         infections_std  = infections/ infections[which.max(mean_conc)],
         positive_std = positive/max(positive)) %>%
  ungroup()

```

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Mean Wastewater Concentration\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Mean Wastewater Concentration\n(Standardized)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Covidestim Estimates\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Covidestim Estimates\n(Standardized)",
       title = "")

```

### Combined Figure

```{r, fig.height=10, fig.width=10}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized)",
       y = "Wastewater Concentrations\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

```

### Compare to Wastewater Over Time

```{r,fig.width=12, fig.height=15}

library(scales)

pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")

names(pal) <- c(#"$observed$", 
                '$P(S_1|untested)$ Centered at Emp. Value',
                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
               "$\\beta$ Centered at Emp. Value"
               )

waste_df <- w_data %>%
  select(biweek, mean_conc_std, name) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc_std=unique(mean_conc_std)) %>%
  mutate(name = gsub("County, MA", "", name))

w_data%>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  filter(version %in% c("v1","v3")) %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup() %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb_std,
                 ymax = exp_cases_ub_std,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df,
                aes(x = date, y =mean_conc_std, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections_std,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data = waste_df,
                aes(x = date, y =mean_conc_std ),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
      facet_grid(name~vlabel, 
                 labeller=labeller(vlabel =as_labeller(
                   TeX, default=label_parsed))) +
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
      theme_bw() +
     theme_c() +
      labs(y = "Value (Standardized)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6)))

```

## Normalization Strategy 2: By Maximum (Separately)

For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Mean Wastewater Concentration\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Mean Wastewater Concentration\n(Standardized)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Covidestim Estimates\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized)",
       y = "Covidestim Estimates\n(Standardized)",
       title = "")
```

### Combined Figure

```{r, fig.height=10, fig.width=10}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized)",
       y = "Wastewater Concentrations\n(Standardized)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

```

### 2023-06-20 reports/priors.Rmd
update rmarkdown files

library(tidyverse)
library(targets)
library(here)

@@

      axis.text.x = element_text(              #axis text
                   size = 12),
      legend.text = element_text(size = 18),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 16, face = "bold", color="white"),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

@@

>>>>

library(tidyverse)
library(targets)
library(here)
library(latex2exp)

@@

      axis.text.x = element_text(              #axis text
                   size = 12),
      legend.text = element_text(size = 14),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 16, face = "bold", color="white"),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

@@

# COVID-19 Trends and Impact Survey

ctis_smoothed <- tar_read(ctis_smoothed,store = here("_targets"))

## $\beta$

```{r beta-smoothing, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$. We compare two approaches to smoothing: cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021) to LOESS smoothing with a span of 0.33."}

ctis_smoothed  %>%
  filter(keep) %>%
  mutate(state=toupper(state)) %>%
  select(date,
         state, 
         imputed_beta,
         beta_estimate_smoothed, 
         beta_estimate_spline_smoothed) %>%
  pivot_longer(contains("beta")) %>%
  mutate(name = case_when(
    name == "beta_estimate_smoothed" ~ "LOESS smoothed",
    name == "beta_estimate_spline_smoothed" ~ "Spline smoothed",
    name == "imputed_beta" ~ "Survey Value"
  )) %>%
  ggplot(aes(x=date, y=value, color = name, 
             alpha = name, linewidth=name)) +
  geom_line() +
  facet_wrap(~state, ncol=4) +
  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
                            "Spline smoothed" = .9, 
                              "Survey Value"=.3),
                     name='') +
  scale_linewidth_manual(values = c("LOESS smoothed" = 1.05, 
                            "Spline smoothed" = 1.05, 
                              "Survey Value"=.5)) +
  scale_color_manual(values=c("#3381FF", "#B58746", "#26900F"),
                     name ='') +
  guides(alpha="none",
         linewidth="none",
         color = guide_legend(override.aes = list(linewidth = 3,
                                                  alpha = c("LOESS smoothed" = .9, 
                            "Spline smoothed" = .9, 
                              "Survey Value"=.3)),
                            nrow=3)) +
  theme_c(legend.position="top") +
  ylim(0,1) +
  scale_x_date(date_breaks="3 months",
               date_labels = "%b %Y") + 
  labs(y = TeX("Survey Estimate of $\\beta$"),
       x= "",
       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $\\beta$"))

## $\Pr(S_1 | \text{untested})$

```{r s-untested-smoothing, fig.cap="The percentage of the population experiencing COVID-19-like illness is taken to be the estimate of $\\Pr(S_1|\\text{untested}). The LOESS smoothed estimate with a span of 0.2 is shown in red."}

ctis_smoothed  %>%
  mutate(state=toupper(state)) %>%
  select(date,
         state, 
         contains("s_untested")) %>%
  pivot_longer(contains("s_untested")) %>%
  mutate(name = case_when(
    name == "s_untested_smoothed" ~ "LOESS smoothed",
    name == "imputed_s_untested" ~ "Survey Value"
  )) %>%
  ggplot(aes(x=date, y=value, color = name, 
             alpha = name, linewidth=name)) +
  geom_line() +
  facet_wrap(~state, ncol=5) +
  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
                              "Survey Value"=.6),
                     name='') +
  scale_linewidth_manual(values = c("LOESS smoothed" = 1.02, 
                              "Survey Value"=.9)) +
  scale_color_manual(values=c("Survey Value" ="black", "LOESS smoothed"="darkred"),
                     name ='') +
  guides(alpha="none",
         linewidth="none",
         color = guide_legend(override.aes = list(
           linewidth = 3,
           alpha = c("LOESS smoothed" = .9, 
                     "Survey Value"=.3)),
                            nrow=3)) +
  theme_c(legend.position="top") +
  scale_x_date(date_breaks="3 months",
               date_labels = "%b %Y") + 
  labs(y = TeX("Survey Estimate of $Pr(S_1|untested)$"),
       x= "",
       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $Pr(S_1|untested)$"))
